---
title: "Describe data"
author: "K. Fleetwood"
date: "`r Sys.Date()`"
output:
  word_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)

## required packages
library(ggplot2)
library(tidyverse)
library(DBI)
library(dbplyr)
library(dplyr)
library(janitor)
library(knitr)
library(arsenal)
library(lubridate)
library(ggh4x) # hacks for ggplot
library(survival)
library(RColorBrewer)

ccu046_folder <- "~/collab/CCU046"
ccu046_02_folder <- file.path(ccu046_folder, "CCU046_02")
pd_folder <- file.path(ccu046_02_folder, "processed_data")
bc_folder <- file.path(ccu046_02_folder, "baseline_characteristics")

r_folder <- file.path(ccu046_folder, "r")
# Additional tableby functions to help with rounding to 5
source(file.path(r_folder, "tableby_functions_19Jan24.R"))  
# Additional functions to help with rounding to 5
source(file.path(r_folder, "output_functions_11Jul24.R")) 
```

```{r load_data, include=FALSE}
flow_out     <- readRDS(file = file.path(pd_folder, "01_flow_out.rds"))
prev_mi_days <- readRDS(file = file.path(pd_folder, "01_prev_mi_days.rds"))
cohort       <- readRDS(file = file.path(pd_folder, "01_cohort.rds"))

st_deviation_no <- c("No acute changes", "Left bundle branch block", "T wave changes only", "Other acute abnormality")

# Alternative labelling for period 4
p4_lab_lvls <- 
  c("Nov 2019 - Feb 2020", "Mar 2020 - Jun 2020", "Jul 2020 - Oct 2020",
    "Nov 2020 - Feb 2021", "Mar 2021 - Jun 2021", "Jul 2021 - Oct 2021", 
    "Nov 2021 - Feb 2022", "Mar 2022 - Jun 2022", "Jul 2022 - Oct 2022", 
    "Nov 2022 - Feb 2023")

cohort <- 
  cohort %>%
  mutate(
    st_deviation = 
      case_when(
        ecg %in% st_deviation_no  ~ "No",
        ecg %in% c("ST segment elevation", "ST segment depression") ~ "Yes",
        is.na(ecg) ~ as.character(NA)
      ), 
    crt_log = log(creatinine),
    troponin_num = as.numeric(peak_troponin),
    admission_to_cardiac_ward = 
      case_when(
        admission_to_cardiac_ward %in% "Died_in_A&E" ~ "No",
        TRUE ~ admission_to_cardiac_ward
      ),
     mi_event_month = floor_date(mi_event_date, "month"),
    smi_bin =
      case_when(
        smi %in% c("Schizophrenia", "Bipolar disorder", "Depression") ~ "Mental illness",
        smi %in% "No SMI" ~ "No mental illness"
      ),
    smi_bin = factor(smi_bin, levels = c("No mental illness", "Mental illness")),
    period_4_lab = paste(substring(period_4, 1, 10), substring(period_4, 12,100)),
    period_4_lab = factor(period_4_lab, p4_lab_lvls)
  )

labels(cohort) <- c(crt_log = "log [Creatinine (micromol/L)]", st_deviation = "ST deviation")
```

# Background
This output includes descriptive tables that will be included in a publication about severe mental illness and mortality following myocardial infarction (project CCU046_02).The output has been created from Myocardial Ischaemia National Audit Project (MINAP) data linked to HES, GDPPR, death records and COVID-19 surveillance data.

# Cohort table
The numbers from this table will be used to make a cohort flow diagram. 
```{r flow}
kable(flow_out, caption = "Number of people in the cohort, rounded to the nearest 5")
```

# Number of people with severe mental illness by type of myocardial infarction
```{r sum_tab}
# Create summary table
nstemi_n <- rnd5(sum(cohort$mi_type %in% "NSTEMI"))
stemi_n  <- rnd5(sum(cohort$mi_type %in% "STEMI"))

sum_tab <- 
  cohort %>% 
  tabyl(smi, mi_type) %>%
  mutate(
    NSTEMI = rnd5(as.numeric(NSTEMI)),
    STEMI = rnd5(as.numeric(STEMI)),
    nstemi_per = 100*NSTEMI/nstemi_n,
    stemi_per  = 100*STEMI/stemi_n,
    NSTEMI = paste0(NSTEMI, " (", format(nstemi_per, digits=1, nsmall=1, trim = TRUE), "%)"), 
    STEMI = paste0(STEMI,  " (", format(stemi_per,  digits=1, nsmall=1, trim = TRUE), "%)")
  ) %>%
  dplyr::select(smi, NSTEMI, STEMI)

kable(sum_tab, caption = "Number of people with severe mental illness by myocardial infarction type, rounded to the nearest 5", col.names = c("Severe mental illness", "NSTEMI", "STEMI"))
```

# Previous MI
Explore different cut-offs for the previous MI variable. Here 0 corresponds to excluding 0 days prior to the MI, i.e. count any previous record of an MI, including records on the day immediately prior to the MI (which I think are likely the same MI; 1 corresponds to excluding the 1 day immediately prior to the MI, etc. 
```{r prev_mi}
kable(prev_mi_days, caption = "Percentage of people with previous MI using different cut-offs for the number of days", col.names = c("Exclude MIs within the x days leading up to the current MI", "NSTEMI", "STEMI"))
```

# NSTEMI: baseline characteristics by severe mental illness
## Key cohorts
```{r nstemi_key, results = "asis"}
# Full cohort
nstemi <- filter(cohort, mi_type %in% "NSTEMI")
nstemi_n <- nstemi %>% nrow()

# Define complete cases flags
nstemi <- 
  nstemi %>%
  mutate(
    cc = !is.na(imd_2019) & !is.na(ethnic_5) & !is.na(cardiogenic_shock) & !is.na(ecg) & !is.na(cardiac_arrest) & 
         !is.na(creatinine) & !is.na(sbp) & !is.na(heart_rate) & !is.na(procode3),
    cc2 = !is.na(imd_2019) & !is.na(ethnic_5) & !is.na(procode3),
    cc_care = cc &  !is.na(angiogram) & !is.na(prev_all) &  !is.na(cardiac_rehab) & !is.na(admission_to_cardiac_ward),
    cc2_care = cc2 &  !is.na(angiogram) & !is.na(prev_all) & !is.na(cardiac_rehab) & !is.na(admission_to_cardiac_ward)
  )

# Full cohort, complete cases
nstemi_cc <- nstemi %>% filter(cc)
nstemi_cc_n <- nstemi_cc %>% nrow()

# Full cohort, complete cases (sensitivity analysis)
nstemi_cc2   <- nstemi %>% filter(cc2)
nstemi_cc2_n <- nstemi_cc2 %>% nrow()

# Surviving 30 days (for analysis of one year mortality amongst people surviving 30 days)
nstemi_30 <- filter(nstemi, !mort_30)
nstemi_30_n <- nstemi_30 %>% nrow()

# Surviving 30 days, complete cases
# Use same variables for complete cases as per full cohort
nstemi_30_cc <- nstemi_30 %>% filter(cc)
nstemi_30_cc_n <- nstemi_30_cc %>% nrow()

# Surviving 30 days, complete cases (sensitivity analysis)
# Use same variables for complete cases as per full cohort
nstemi_30_cc2 <- nstemi_30 %>% filter(cc2)
nstemi_30_cc2_n <- nstemi_30_cc2 %>% nrow()

# Discharged home
nstemi_home <- filter(nstemi, discharge_destination %in% "Home")
nstemi_home_n <- nstemi_home %>% nrow()

# Discharged home, complete cases
nstemi_home_cc <- nstemi_home %>% filter(cc_care)
nstemi_home_cc_n <- nstemi_home_cc %>% nrow()

# Discharged home, complete cases (sensitivity analysis)
nstemi_home_cc2 <- nstemi_home %>% filter(cc2_care)
nstemi_home_cc2_n <- nstemi_home_cc2 %>% nrow()

# Create table
key_cohorts <- 
  data.frame(
    cohort = c("Full cohort", "Survived more than 30 days", "Discharged home"),
    n = c(nstemi_n,       nstemi_30_n,       nstemi_home_n),
    n_cc = c(nstemi_cc_n,    nstemi_30_cc_n,    nstemi_home_cc_n),
    n_cc2 = c(nstemi_cc2_n,    nstemi_30_cc2_n,    nstemi_home_cc2_n)
  ) %>%
  mutate(
    n_rnd = rnd5(n),
    #
    n_cc_rnd = rnd5(n_cc),
    per_cc = round(100*n_cc_rnd/n_rnd, digits = 1),
    n_cc_rnd_fmt = paste0(format(n_cc_rnd, digits=6), " (", per_cc, "%)"),
    #
    n_cc2_rnd = rnd5(n_cc2),
    per_cc2 = round(100*n_cc2_rnd/n_rnd, digits = 1),
    n_cc2_rnd_fmt = paste0(format(n_cc2_rnd, digits=6), " (", per_cc2, "%)")
  )

kable(dplyr::select(key_cohorts, cohort, n_rnd, n_cc_rnd_fmt, n_cc2_rnd_fmt), caption = "NSTEMI: Key cohorts", col.names = c("Cohort", "N", "Complete cases", "Complete cases (SA)"))
```

For the full cohort and the cohort who survived more than 30 days, for the main analyses, the complete cases are based on IMD decile, ethnicity, cardiogenic shock, ECG determining treatment, cardiac arrest, creatinine, SBP, heart rate and hospital.

In a sensitivity analysis (SA) excluding the MI characteristics, the complete cases analyses are based on IMD decile, ethnicity and hosptial only.

For the cohort of people discharged home complete cases are additionally based on the processes of care (receipt of angiogram, admission to cardiac ward, receipt of secondary prevention medication and cardiac rehabilitation).

```{r nstemi_smi, results = "asis"}
nstemi_smi               <- smi_fun(nstemi)
nstemi_cc_smi            <- smi_fun(nstemi_cc)
nstemi_cc2_smi            <- smi_fun(nstemi_cc2)

nstemi_30_smi            <- smi_fun(nstemi_30)
nstemi_30_cc_smi         <- smi_fun(nstemi_30_cc)
nstemi_30_cc2_smi         <- smi_fun(nstemi_30_cc2)

nstemi_home_smi          <- smi_fun(nstemi_home)
nstemi_home_cc_smi       <- smi_fun(nstemi_home_cc)
nstemi_home_cc2_smi       <- smi_fun(nstemi_home_cc2)

# NSTEMI cohorts by SMI
nstemi_cohorts_smi <- 
  rbind(nstemi_smi, nstemi_cc_smi, nstemi_cc2_smi,
        nstemi_30_smi, nstemi_30_cc_smi, nstemi_30_cc2_smi,
        nstemi_home_smi, nstemi_home_cc_smi, nstemi_home_cc2_smi)

colnames(nstemi_cohorts_smi) <- c("Schizophrenia", "Bipolar disorder", "Depression", "No SMI")

nstemi_cohorts_smi <- 
  as.data.frame(nstemi_cohorts_smi) %>%
  mutate(
    Cohort = rep(c("Full cohort", "Survived more than 30 days", "Discharged home"), each = 3),
    Including = rep(c("All", "CC", "CC (SA)"), times = 3)
  ) %>%
  dplyr::select(Cohort, Including, Schizophrenia, `Bipolar disorder`, Depression, `No SMI`)

kable(nstemi_cohorts_smi, caption = "NSTEMI cohorts by SMI", row.names = FALSE)
```
CC: complete cases, CC: complete cases (sensitivity analysis)

## Full cohort
### Including missing data
```{r nstemi_base, results = "asis"}
tab1_cntrls <- 
  tableby.control(
    numeric.stats = c("meansd" ,"medianrange_kf" , "n_per_miss_5"),
    cat.stats = c("countpct_5"), cat.simplify = FALSE
  )

# Baseline characteristics table based on full NSTEMI cohort
tab1_nstemi <- 
  tableby(
    smi ~
      age_mi + sex + includeNA(imd_2019, label = "Missing") + includeNA(ethnic_5, label = "Missing") + 
      mi_event_year + mi_wd + mi_on + procode3_miss + 
      comorb_angina + comorb_mi + hx_pci + comorb_hf +  comorb_diab + comorb_crf + comorb_cevd + 
      includeNA(cardiogenic_shock, label = "Missing") + 
      includeNA(ecg, label = "Missing") + 
      includeNA(st_deviation, label = "Missing") + 
      includeNA(cardiac_arrest, label = "Missing") + 
      includeNA(cardiac_enzymes_elevated, label = "Missing") +
      creatinine + crt_log + sbp + heart_rate +  troponin_num + covid + 
      includeNA(bmi_cat) + includeNA(smoking_cat) +
      mort_30 + mort_yr + cod_cvd_yr +
      includeNA(discharge_destination),
    data = filter(cohort, mi_type %in% "NSTEMI"),
    control = tab1_cntrls, 
    digits = 1, 
    stats.labels = 
      list(meansd = 'Mean (SD)',
           medianrange_kf = 'Median [IQR]',
           n_per_miss = 'Missing',
           n_per_miss_5 = 'Missing',
           Nmiss = 'Missing')
  )
tab1_nstemi$tables$smi$y$stats <- 5*round(tab1_nstemi$tables$smi$y$stats/5)

summary(tab1_nstemi, title = "Baseline characteristics for the NSTEMI cohort, counts rounded to the nearest 5")

write.csv(summary(tab1_nstemi, text = NULL), file = file.path(bc_folder, "02_nstemi_full_baseline.csv"))

# Top 10 causes of death within 30 days, for people who died within 30 days
tab1_nstemi_cod30 <- 
  tableby(
    smi ~ cod_top10_mort30_nstemi,
    data = filter(cohort, mi_type %in% "NSTEMI", mort_30),
    control = tab1_cntrls, 
    digits = 1, 
    stats.labels = 
      list(meansd = 'Mean (SD)',
           medianrange_kf = 'Median [IQR]',
           n_per_miss = 'Missing',
           n_per_miss_5 = 'Missing',
           Nmiss = 'Missing')
  )
tab1_nstemi_cod30$tables$smi$y$stats <- 5*round(tab1_nstemi_cod30$tables$smi$y$stats/5)

summary(tab1_nstemi_cod30, title = "Top 10 causes of mortality within 30 days, NSTEMI cohort, counts rounded to the nearest 5")

# Top 10 causes of death within a year, for people who died within a year
tab1_nstemi_cod1 <- 
  tableby(
    smi ~ cod_top10_mort_yr_nstemi,
    data = filter(cohort, mi_type %in% "NSTEMI", mort_yr),
    control = tab1_cntrls, 
    digits = 1, 
    stats.labels = 
      list(meansd = 'Mean (SD)',
           medianrange_kf = 'Median [IQR]',
           n_per_miss = 'Missing',
           n_per_miss_5 = 'Missing',
           Nmiss = 'Missing')
  )
tab1_nstemi_cod1$tables$smi$y$stats <- 5*round(tab1_nstemi_cod1$tables$smi$y$stats/5)

summary(tab1_nstemi_cod1, title = "Top 10 causes of mortality within one year, NSTEMI cohort, counts rounded to the nearest 5")
```

### Complete cases
```{r nstemi_cc_base, results = "asis"}
tab1_nstemi_cc <- 
  tableby(
    smi ~
      age_mi + sex + imd_2019 + ethnic_5 + 
      mi_event_year + mi_wd + mi_on + procode3_miss + 
      comorb_angina + comorb_mi + hx_pci + comorb_hf +  comorb_diab + comorb_crf + comorb_cevd + 
      cardiogenic_shock + ecg + st_deviation + cardiac_arrest + includeNA(cardiac_enzymes_elevated)+
      creatinine + crt_log + sbp + heart_rate + covid + 
      includeNA(bmi_cat) + includeNA(smoking_cat) +
      mort_30 + mort_yr + cod_cvd_yr +
      includeNA(discharge_destination),
    data = nstemi_cc,
    control = tab1_cntrls, 
    digits = 1, 
    stats.labels = 
      list(meansd = 'Mean (SD)',
           medianrange_kf = 'Median [IQR]',
           n_per_miss = 'Missing',
           n_per_miss_5 = 'Missing',
           Nmiss = 'Missing')
  )
tab1_nstemi_cc$tables$smi$y$stats <- 5*round(tab1_nstemi_cc$tables$smi$y$stats/5)

summary(tab1_nstemi_cc, title = "Baseline characteristics for the NSTEMI cohort (complete cases), counts rounded to the nearest 5")

write.csv(summary(tab1_nstemi_cc, text = NULL), file = file.path(bc_folder, "02_nstemi_cc_baseline.csv"))
```

### Complete cases (sensitivity analysis)
```{r nstemi_cc2_base, results = "asis"}
tab1_nstemi_cc2 <- 
  tableby(
    smi ~
      age_mi + sex + imd_2019 + ethnic_5 + 
      mi_event_year + mi_wd + mi_on + procode3_miss + 
      comorb_angina + comorb_mi + hx_pci + comorb_hf +  comorb_diab + comorb_crf + comorb_cevd + 
      mort_30 + mort_yr + cod_cvd_yr +
      includeNA(discharge_destination),
    data = nstemi_cc2,
    control = tab1_cntrls, 
    digits = 1, 
    stats.labels = 
      list(meansd = 'Mean (SD)',
           medianrange_kf = 'Median [IQR]',
           n_per_miss = 'Missing',
           n_per_miss_5 = 'Missing',
           Nmiss = 'Missing')
  )
tab1_nstemi_cc2$tables$smi$y$stats <- 5*round(tab1_nstemi_cc2$tables$smi$y$stats/5)

summary(tab1_nstemi_cc2, title = "Baseline characteristics for the NSTEMI cohort (complete cases, sensitivity analysis), counts rounded to the nearest 5")

write.csv(summary(tab1_nstemi_cc2, text = NULL), file = file.path(bc_folder, "02_nstemi_cc_sa_baseline.csv"))
```

## People surviving 30 days 
### Including missing data
```{r nstemi_30_base, results = "asis"}

tab1_nstemi_30 <- 
  tableby(
    smi ~
      age_mi + sex + 
      includeNA(imd_2019, label = "Missing") + 
      includeNA(ethnic_5, label = "Missing") + 
      mi_event_year + mi_wd + mi_on + procode3_miss + 
      comorb_angina + comorb_mi + hx_pci + comorb_hf +  comorb_diab + comorb_crf + comorb_cevd + 
      includeNA(cardiogenic_shock, label = "Missing") + 
      includeNA(ecg, label = "Missing") + 
      includeNA(st_deviation, label = "Missing") + 
      includeNA(cardiac_arrest, label = "Missing") + 
      includeNA(cardiac_enzymes_elevated, label = "Missing") +
      creatinine + crt_log + sbp + heart_rate + covid + 
      includeNA(bmi_cat) + includeNA(smoking_cat) +
      mort_yr + cod_cvd_yr +
      includeNA(discharge_destination),
    data = nstemi_30,
    control = tab1_cntrls, 
    digits = 1, 
    stats.labels = 
      list(meansd = 'Mean (SD)',
           medianrange_kf = 'Median [IQR]',
           n_per_miss = 'Missing',
           n_per_miss_5 = 'Missing',
           Nmiss = 'Missing')
  )
tab1_nstemi_30$tables$smi$y$stats <- 5*round(tab1_nstemi_30$tables$smi$y$stats/5)

summary(tab1_nstemi_30, title = "Baseline characteristics for the NSTEMI cohort, people who survived more than 30 days, counts rounded to the nearest 5")
```

### Complete cases
```{r nstemi_30_cc_base, results = "asis"}

tab1_nstemi_30_cc <- 
  tableby(
    smi ~
      age_mi + sex + imd_2019 + ethnic_5 + 
      mi_event_year + mi_wd + mi_on + procode3_miss + 
      comorb_angina + comorb_mi + hx_pci + comorb_hf +  comorb_diab + comorb_crf + comorb_cevd + 
      cardiogenic_shock + ecg + st_deviation + cardiac_arrest + includeNA(cardiac_enzymes_elevated)+
      creatinine + crt_log + sbp + heart_rate + covid + 
      includeNA(bmi_cat) + includeNA(smoking_cat) +
      mort_yr + cod_cvd_yr +
      includeNA(discharge_destination),
    data = nstemi_30_cc,
    control = tab1_cntrls, 
    digits = 1, 
    stats.labels = 
      list(meansd = 'Mean (SD)',
           medianrange_kf = 'Median [IQR]',
           n_per_miss = 'Missing',
           n_per_miss_5 = 'Missing',
           Nmiss = 'Missing')
  )
tab1_nstemi_30_cc$tables$smi$y$stats <- 5*round(tab1_nstemi_30_cc$tables$smi$y$stats/5)

summary(tab1_nstemi_30_cc, title = "Baseline characteristics for the NSTEMI cohort, people who survived more than 30 days (complete cases), counts rounded to the nearest 5")
```

### Complete cases (sensitivity analysis)
```{r nstemi_30_cc2_base, results = "asis"}

tab1_nstemi_30_cc2 <- 
  tableby(
    smi ~
      age_mi + sex + 
      imd_2019 + 
      ethnic_5 + 
      mi_event_year + mi_wd + mi_on + procode3_miss + 
      comorb_angina + comorb_mi + hx_pci + comorb_hf +  comorb_diab + comorb_crf + comorb_cevd + 
      mort_yr + cod_cvd_yr +
      includeNA(discharge_destination),
    data = nstemi_30_cc2,
    control = tab1_cntrls, 
    digits = 1, 
    stats.labels = 
      list(meansd = 'Mean (SD)',
           medianrange_kf = 'Median [IQR]',
           n_per_miss = 'Missing',
           n_per_miss_5 = 'Missing',
           Nmiss = 'Missing')
  )

tab1_nstemi_30_cc2$tables$smi$y$stats <- 5*round(tab1_nstemi_30_cc2$tables$smi$y$stats/5)

summary(tab1_nstemi_30_cc2, title = "Baseline characteristics for the NSTEMI cohort, people who survived more than 30 days (complete cases, senstivity analysis), counts rounded to the nearest 5")
```

## People discharged home
### Including missing data
```{r nstemi_home_base, results = "asis"}

tab1_nstemi_home <- 
  tableby(
    smi ~
      age_mi + sex + includeNA(imd_2019, label = "Missing") + includeNA(ethnic_5, label = "Missing") + 
      mi_event_year + mi_wd + mi_on + procode3_miss + 
      comorb_angina + comorb_mi + hx_pci + comorb_hf +  comorb_diab + comorb_crf + comorb_cevd + 
      includeNA(cardiogenic_shock, label = "Missing") + 
      includeNA(ecg, label = "Missing") + 
      includeNA(st_deviation, label = "Missing") + 
      includeNA(cardiac_arrest, label = "Missing") + 
      includeNA(cardiac_enzymes_elevated, label = "Missing") +
      creatinine + crt_log + sbp + heart_rate +  covid + 
      includeNA(bmi_cat) + includeNA(smoking_cat) +
      includeNA(angiogram) + 
      includeNA(admission_to_cardiac_ward) + 
      includeNA(prev_all) + 
      includeNA(cardiac_rehab) +
      mort_yr + cod_cvd_yr,
    data = nstemi_home,
    control = tab1_cntrls, 
    digits = 1, 
    stats.labels = 
      list(meansd = 'Mean (SD)',
           medianrange_kf = 'Median [IQR]',
           n_per_miss = 'Missing',
           n_per_miss_5 = 'Missing',
           Nmiss = 'Missing')
  )
tab1_nstemi_home$tables$smi$y$stats <- 5*round(tab1_nstemi_home$tables$smi$y$stats/5)

summary(tab1_nstemi_home, title = "Baseline characteristics for the NSTEMI cohort, people discharged home, counts rounded to the nearest 5")

write.csv(summary(tab1_nstemi_home, text = NULL), file = file.path(bc_folder, "02_nstemi_home_full_baseline.csv"))
```

### Complete cases
```{r nstemi_home_cc_base, results = "asis"}

tab1_nstemi_home_cc <- 
  tableby(
    smi ~
      age_mi + sex + imd_2019 + ethnic_5 + 
      mi_event_year + mi_wd + mi_on + procode3_miss + 
      comorb_angina + comorb_mi + hx_pci + comorb_hf +  comorb_diab + comorb_crf + comorb_cevd + 
      cardiogenic_shock + ecg + st_deviation + cardiac_arrest + includeNA(cardiac_enzymes_elevated)+
      creatinine + crt_log + sbp + heart_rate +  covid + 
      includeNA(bmi_cat) + includeNA(smoking_cat) +
      angiogram + admission_to_cardiac_ward + prev_all + cardiac_rehab +
      mort_yr + cod_cvd_yr,
    data = nstemi_home_cc,
    control = tab1_cntrls, 
    digits = 1, 
    stats.labels = 
      list(meansd = 'Mean (SD)',
           medianrange_kf = 'Median [IQR]',
           n_per_miss = 'Missing',
           n_per_miss_5 = 'Missing',
           Nmiss = 'Missing')
  )
tab1_nstemi_home_cc$tables$smi$y$stats <- 5*round(tab1_nstemi_home_cc$tables$smi$y$stats/5)

summary(tab1_nstemi_home_cc, title = "Baseline characteristics for the NSTEMI cohort, people discharged home (complete cases), counts rounded to the nearest 5")

write.csv(summary(tab1_nstemi_home_cc, text = NULL), file = file.path(bc_folder, "02_nstemi_home_cc_baseline.csv"))
```

### Complete cases (sensitivity analysis)
```{r nstemi_home_cc2_base, results = "asis"}

tab1_nstemi_home_cc2 <- 
  tableby(
    smi ~
      age_mi + sex + imd_2019 + ethnic_5 + 
      mi_event_year + mi_wd + mi_on + procode3_miss + 
      comorb_angina + comorb_mi + hx_pci + comorb_hf +  comorb_diab +  comorb_crf + comorb_cevd + 
      angiogram + admission_to_cardiac_ward + prev_all + cardiac_rehab +
      mort_yr + cod_cvd_yr,
    data = nstemi_home_cc2,
    control = tab1_cntrls, 
    digits = 1, 
    stats.labels = 
      list(meansd = 'Mean (SD)',
           medianrange_kf = 'Median [IQR]',
           n_per_miss = 'Missing',
           n_per_miss_5 = 'Missing',
           Nmiss = 'Missing')
  )
tab1_nstemi_home_cc2$tables$smi$y$stats <- 5*round(tab1_nstemi_home_cc2$tables$smi$y$stats/5)

summary(tab1_nstemi_home_cc2, title = "Baseline characteristics for the NSTEMI cohort, people discharged home (complete cases, sensitivity analysis), counts rounded to the nearest 5")

write.csv(summary(tab1_nstemi_home_cc2, text = NULL), file = file.path(bc_folder, "02_nstemi_home_cc_sa_baseline.csv"))
```

# NSTEMI: descriptive plots
## Summary of outcomes by month
```{r out_month, results = "asis"}

# Count of NSTEMI by month
nstemi_month_adm <- 
  nstemi %>%
  tabyl(mi_event_month) %>%
  mutate(
    outcome = "MI admissions (N)",
    n_per = rnd5(n)
  ) %>%
  dplyr::select(outcome, mi_event_month, n_per)

# 30-day mortality
nstemi_month_30day <-
  nstemi %>%
  mutate(
    mort_30 = ifelse(mort_30, "Yes", "No")
  ) %>%
  tabyl(mi_event_month, mort_30) %>%
  mutate(
    outcome = "30-day mortality (%)",
    total = No + Yes,
    n_per = 100*rnd5(Yes)/rnd5(total)
  ) %>%
  dplyr::select(outcome, mi_event_month, n_per)

# 1-year mortality
nstemi_month_1yr <-
  nstemi %>%
  mutate(
    mort_yr = ifelse(mort_yr, "Yes", "No")
  ) %>%
  tabyl(mi_event_month, mort_yr) %>%
  mutate(
    outcome = "1-year mortality (%)",
    total = No + Yes,
    n_per = 100*rnd5(Yes)/rnd5(total)
  ) %>%
  dplyr::select(outcome, mi_event_month, n_per)

# 1-year CVD mortality
nstemi_month_cvd_1yr <-
  nstemi %>%
  mutate(
    mort_cvd_yr = ifelse(mort_cvd_yr, "Yes", "No")
  ) %>%
  tabyl(mi_event_month, mort_cvd_yr) %>%
  mutate(
    outcome = "1-year CVD mortality (%)",
    total = No + Yes,
    n_per = 100*rnd5(Yes)/rnd5(total)
  ) %>%
  dplyr::select(outcome, mi_event_month, n_per)

# 1-year mortality amongst people who survived 30 days
nstemi_month_1yr_30 <-
  nstemi_30 %>%
  mutate(
    mort_yr = ifelse(mort_yr, "Yes", "No")
  ) %>%
  tabyl(mi_event_month, mort_yr) %>%
  mutate(
    outcome = "1-year mortality (amongst people surviving 30 days) (%)",
    total = No + Yes,
    n_per = 100*rnd5(Yes)/rnd5(total)
  ) %>%
  dplyr::select(outcome, mi_event_month, n_per)

# 1-year mortality amongst people discharged home
nstemi_month_1yr_home <-
  nstemi_home %>%
  mutate(
    mort_yr = ifelse(mort_yr, "Yes", "No")
  ) %>%
  tabyl(mi_event_month, mort_yr) %>%
  mutate(
    outcome = "1-year mortality (amongst people discharged home) (%)",
    total = No + Yes,
    n_per = 100*rnd5(Yes)/rnd5(total)
  ) %>%
  dplyr::select(outcome, mi_event_month, n_per)
  
outcome_lvls <-
  c("MI admissions (N)", "30-day mortality (%)", "1-year mortality (%)", 
    "1-year CVD mortality (%)", 
    "1-year mortality (amongst people surviving 30 days) (%)",
    "1-year mortality (amongst people discharged home) (%)")

# Collate outcomes by month
nstemi_month <- 
  bind_rows(
    nstemi_month_adm,
    nstemi_month_30day,
    nstemi_month_1yr,
    nstemi_month_cvd_1yr,
    nstemi_month_1yr_30,
    nstemi_month_1yr_home
  ) %>%
  mutate(
    outcome = factor(outcome, levels = outcome_lvls)
  )

# Save outcomes by month
write.csv(nstemi_month, file.path(bc_folder, "02_nstemi_month.csv"), row.names = FALSE)

nstemi_month_plot <- 
ggplot(nstemi_month, aes(x = mi_event_month, y = n_per)) + 
  geom_line() + 
  facet_wrap(
    vars(outcome),
    ncol = 1,
    scales = "free_y",
    strip.position = "top"
  ) +
  scale_x_date(date_labels = "%h %Y", expand = c(0,0)) + 
  theme_bw() + 
  theme(strip.background = element_rect(fill = "white"), panel.spacing = unit(0 , 'lines')) + 
  labs(x = "Month", y = "")

nstemi_month_plot
```

## Summary of outcomes by 4-month period
```{r out_period, results = "asis"}

# Count of NSTEMI by period and SMI
nstemi_per_adm <- 
  nstemi %>%
  group_by(period_4, smi) %>%
  summarise(
    n = n()
  ) %>%
  mutate(
    outcome = "MI admissions (N)",
    n_per = rnd5(n)
  ) %>%
  dplyr::select(outcome, period_4, smi, n_per)

# 30-day mortality
nstemi_per_30day <-
  nstemi %>%
  group_by(period_4, smi) %>%
  summarise(
    outcome = "30-day mortality (%)",
    total = n(),
    Yes = sum(mort_30),
    n_per = 100*Yes/total
  ) %>%
  dplyr::select(outcome, period_4, smi, n_per)

# 1-year mortality
nstemi_per_1yr <-
  nstemi %>%
  group_by(period_4, smi) %>%
  summarise(
    outcome = "1-year mortality (%)",
    total = n(),
    Yes = sum(mort_yr),
    n_per = 100*Yes/total
  ) %>%
  dplyr::select(outcome, period_4, smi, n_per)

# 1-year CVD mortality
nstemi_per_cvd_1yr <-
  nstemi %>%
  group_by(period_4, smi) %>%
  summarise(
    outcome = "1-year CVD mortality (%)",
    total = n(),
    Yes = sum(mort_cvd_yr),
    n_per = 100*Yes/total
  ) %>%
  dplyr::select(outcome, period_4, smi, n_per)

# 1-year mortality (amongst people surviving 30 days)
nstemi_per_1yr_30 <- 
  nstemi_30 %>%
  group_by(period_4, smi) %>%
  summarise(
    outcome = "1-year mortality (amongst people surviving 30 days) (%)",
    total = n(),
    Yes = sum(mort_yr),
    n_per = 100*Yes/total
  ) %>%
  dplyr::select(outcome, period_4, smi, n_per)

# 1-year mortality (amongst people discharged home)
nstemi_per_1yr_home <- 
  nstemi_home %>%
  group_by(period_4, smi) %>%
  summarise(
    outcome = "1-year mortality (amongst people discharged home) (%)",
    total = n(),
    Yes = sum(mort_yr),
    n_per = 100*Yes/total
  ) %>%
  dplyr::select(outcome, period_4, smi, n_per)
  
# Collate outcomes by month
nstemi_per <- 
  bind_rows(
    nstemi_per_adm,
    nstemi_per_30day,
    nstemi_per_1yr,
    nstemi_per_cvd_1yr,
    nstemi_per_1yr_30,
    nstemi_per_1yr_home
  ) %>%
  mutate(
    outcome = factor(outcome, levels = outcome_lvls),
    period_4_start = paste("1", substring(period_4, 1, 8)),
    period_4_date = as.Date(period_4_start, format = "%d %b %Y"),
  )

# Save outcomes by month
write.csv(nstemi_per, file.path(bc_folder, "02_nstemi_period_smi.csv"), row.names = FALSE)

nstemi_per_plot <- 
ggplot(nstemi_per, aes(x = period_4_date, y = n_per, color = smi)) + 
  geom_line() + 
  facet_wrap(
    vars(outcome),
    ncol = 1,
    scales = "free_y",
    strip.position = "top"
  ) +
  scale_x_date(date_labels = "%h %Y", expand = c(0,0)) + 
  theme_bw() + 
  theme(strip.background = element_rect(fill = "white"), panel.spacing = unit(0 , 'lines')) + 
  labs(x = "Month", y = "")

nstemi_per_plot
```

## Summary of outcomes by 4-month period (binary SMI)
```{r out_period_bin, results = "asis"}

# Count of NSTEMI by period and SMI
nstemi_per_bin_adm <- 
  nstemi %>%
  group_by(period_4, smi_bin) %>%
  summarise(
    n = n()
  ) %>%
  mutate(
    outcome = "MI admissions (N)",
    n_per = rnd5(n)
  ) %>%
  dplyr::select(outcome, period_4, smi_bin, n_per)

# 30-day mortality
nstemi_per_bin_30day <-
  nstemi %>%
  group_by(period_4, smi_bin) %>%
  summarise(
    outcome = "30-day mortality (%)",
    total = n(),
    Yes = sum(mort_30),
    n_per = 100*Yes/total,
    n_per_low = 100*prop.test(Yes, total)$conf.int[1],
    n_per_upp = 100*prop.test(Yes, total)$conf.int[2]
  ) %>%
  dplyr::select(outcome, period_4, smi_bin, n_per, n_per_low, n_per_upp)

# 1-year mortality
nstemi_per_bin_1yr <-
  nstemi %>%
  group_by(period_4, smi_bin) %>%
  summarise(
    outcome = "1-year mortality (%)",
    total = n(),
    Yes = sum(mort_yr),
    n_per = 100*Yes/total,
    n_per_low = 100*prop.test(Yes, total)$conf.int[1],
    n_per_upp = 100*prop.test(Yes, total)$conf.int[2]
  ) %>%
  dplyr::select(outcome, period_4, smi_bin, n_per, n_per_low, n_per_upp)

# 1-year CVD mortality
nstemi_per_bin_cvd_1yr <-
  nstemi %>%
  group_by(period_4, smi_bin) %>%
  summarise(
    outcome = "1-year CVD mortality (%)",
    total = n(),
    Yes = sum(mort_cvd_yr),
    n_per = 100*Yes/total,
    n_per_low = 100*prop.test(Yes, total)$conf.int[1],
    n_per_upp = 100*prop.test(Yes, total)$conf.int[2]
  ) %>%
  dplyr::select(outcome, period_4, smi_bin, n_per, n_per_low, n_per_upp)

# 1-year mortality (amongst people surviving 30 days)
nstemi_per_bin_1yr_30 <-
  nstemi_30 %>%
  group_by(period_4, smi_bin) %>%
  summarise(
    outcome = "1-year mortality (amongst people surviving 30 days) (%)",
    total = n(),
    Yes = sum(mort_yr),
    n_per = 100*Yes/total,
    n_per_low = 100*prop.test(Yes, total)$conf.int[1],
    n_per_upp = 100*prop.test(Yes, total)$conf.int[2]
  ) %>%
  dplyr::select(outcome, period_4, smi_bin, n_per, n_per_low, n_per_upp)

# 1-year mortality (amongst people discharged home)
nstemi_per_bin_1yr_home <-
  nstemi_home %>%
  group_by(period_4, smi_bin) %>%
  summarise(
    outcome = "1-year mortality (amongst people discharged home) (%)",
    total = n(),
    Yes = sum(mort_yr),
    n_per = 100*Yes/total,
    n_per_low = 100*prop.test(Yes, total)$conf.int[1],
    n_per_upp = 100*prop.test(Yes, total)$conf.int[2]
  ) %>%
  dplyr::select(outcome, period_4, smi_bin, n_per, n_per_low, n_per_upp)

# Collate outcomes by month
nstemi_per_bin <- 
  bind_rows(
    nstemi_per_bin_adm,
    nstemi_per_bin_30day,
    nstemi_per_bin_1yr,
    nstemi_per_bin_cvd_1yr,
    nstemi_per_bin_1yr_30,
    nstemi_per_bin_1yr_home
  ) %>%
  mutate(
    outcome = factor(outcome, levels = outcome_lvls),
    period_4_start = paste("1", substring(period_4, 1, 8)),
    period_4_date = as.Date(period_4_start, format = "%d %b %Y"),
  )

# Save outcomes by month
write.csv(nstemi_per_bin, file.path(bc_folder, "02_nstemi_period_smi_bin.csv"), row.names = FALSE)

nstemi_per_bin_plot <- 
ggplot(nstemi_per_bin, aes(x = period_4_date, y = n_per, color = smi_bin)) + 
  geom_line() + 
  facet_wrap(
    vars(outcome),
    ncol = 1,
    scales = "free_y",
    strip.position = "top"
  ) +
  scale_x_date(date_labels = "%h %Y", expand = c(0,0)) + 
  theme_bw() + 
  theme(strip.background = element_rect(fill = "white"), panel.spacing = unit(0 , 'lines')) + 
  labs(x = "Month", y = "")

nstemi_per_bin_plot
```

## Kaplan Meier plot by SMI
```{r nstemi_km_smi, results = "asis"}

# All-cause mortality ---------------------------------------------------------
nstemi_smi_sf <- survfit(Surv(mort_yr_days, mort_yr) ~ smi, data = nstemi) # Create survfit object

# Kaplan-Meier plot
plot(nstemi_smi_sf, col = 1:4, xlab = "Time (days)", ylab = "Survival", ylim = c(0.8, 1))
legend("topright", legend = levels(cohort$smi), col = 1:4, lty =1)

pdf(file.path(bc_folder, "02_nstemi_km_mort_smi.pdf"), height = 7, width = 7)
plot(nstemi_smi_sf, col = 1:4, xlab = "Time (days)", ylab = "Survival", ylim = c(0.8, 1))
legend("topright", legend = levels(cohort$smi), col = 1:4, lty =1)
dev.off()

nstemi_smi_sf_out <- as.df.sf(nstemi_smi_sf)
write.csv(nstemi_smi_sf_out, file.path(bc_folder, "02_nstemi_smi_sf.csv"))

# CVD-mortality ---------------------------------------------------------------

# Survfit object
nstemi_cvd_smi_sf <- survfit(Surv(mort_yr_days, mort_cvd_yr) ~ smi, data = nstemi) # Create survfit object

# Kaplan meier plot
plot(nstemi_cvd_smi_sf, col = 1:4, xlab = "Time (days)", ylab = "Survival", ylim = c(0.9, 1))
legend("topright", legend = levels(cohort$smi), col = 1:4, lty =1)

pdf(file.path(bc_folder, "02_nstemi_km_cvd_mort_smi.pdf"), height = 7, width = 7)
plot(nstemi_cvd_smi_sf, col = 1:4, xlab = "Time (days)", ylab = "Survival", ylim = c(0.9, 1))
legend("topright", legend = levels(cohort$smi), col = 1:4, lty =1)
dev.off()

nstemi_cvd_smi_sf_out <- as.df.sf(nstemi_cvd_smi_sf)
write.csv(nstemi_cvd_smi_sf_out, file.path(bc_folder, "02_nstemi_cvd_smi_sf.csv"))
```

## Kaplan Meier plot by time period
```{r nstemi_km_per, results = "asis"}

# All-cause mortality ---------------------------------------------------------
nstemi_per_sf <- survfit(Surv(mort_yr_days, mort_yr) ~ period_4_lab, data = nstemi) # Create survfit object

# Kaplan meier plot
plot(nstemi_per_sf, col = brewer.pal(n = 10, name = "Paired"), xlab = "Time (days)", ylab = "Survival", ylim = c(0.85, 1))
legend("topright", legend = levels(cohort$period_4_lab), col = brewer.pal(n = 10, name = "Paired"), lty =1, cex = 0.5)

pdf(file.path(bc_folder, "02_nstemi_km_mort_per.pdf"), height = 7, width = 7)
plot(nstemi_per_sf, col = brewer.pal(n = 10, name = "Paired"), xlab = "Time (days)", ylab = "Survival", ylim = c(0.85, 1))
legend("topright", legend = levels(cohort$period_4_lab), col = brewer.pal(n = 10, name = "Paired"), lty =1, cex = 0.8)
dev.off()

nstemi_per_sf_out <- as.df.sf(nstemi_per_sf)
write.csv(nstemi_per_sf_out, file.path(bc_folder, "02_nstemi_per_sf.csv"))

# CVD-mortality ----------------------------------
nstemi_per_cvd_sf <- survfit(Surv(mort_yr_days, mort_cvd_yr)  ~ period_4_lab, data = nstemi) # Create survfit object

# Kaplan meier plot
plot(nstemi_per_cvd_sf, col = brewer.pal(n = 10, name = "Paired"), xlab = "Time (days)", ylab = "Survival", ylim = c(0.92, 1))
legend("topright", legend = levels(cohort$period_4_lab), col = brewer.pal(n = 10, name = "Paired"), lty =1, cex = 0.5)

pdf(file.path(bc_folder, "02_nstemi_km_mort_cvd_per.pdf"), height = 7, width = 7)
plot(nstemi_per_cvd_sf, col = brewer.pal(n = 10, name = "Paired"), xlab = "Time (days)", ylab = "Survival", ylim = c(0.92, 1))
legend("topright", legend = levels(cohort$period_4_lab), col = brewer.pal(n = 10, name = "Paired"), lty =1, cex = 0.5)
dev.off()
```
